---
title: "Final version of NMDS"
author: "Caroline Owens"
date: "4/22/2022"
output: 
  html_document: 
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r packages}
library(vegan)
library(lubridate)
library(tidyverse)
library(ggrepel)
```

```{r themes, include= FALSE}
th = theme_set(theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)))
```

# Setup

```{r load data}
inci_data = read_csv("cleaned_data/combined_unsummed_etraps.csv") %>% 
  mutate(sample_date = as_date(sample_date, format = "%d-%b-%y"))

lake_data = read_csv("cleaned_data/emerg_model_data.csv")%>% 
  select(c(fish:trap, date_off, peaktemp, dailytemp, DSI)) %>% 
  rename(sample_date = date) %>% 
  mutate(date_off = as_date(date_off))

full_data = lake_data %>% 
  right_join(inci_data, by = c("fish", "block", "lake", "sample_date", "trap")) %>% 
  mutate(index = paste0("trap_",1:nrow(.)), .before = fish) %>% 
  column_to_rownames("index") %>% 
  relocate(data_flag, .before = fish)

rm(lake_data, inci_data)

#write_csv(full_data, "cleaned_data/full_data.csv")
```

```{r reformat data}
# Pool by lake/year; reduce unwanted dimensions

sp_ly = full_data %>% 
  group_by(lake,year) %>% 
  summarize(across(.cols = dip_chir_lg:ephem, .fns = sum)) %>% 
  mutate(id = paste(lake,year, sep = "_"), .keep = "unused") %>% 
  column_to_rownames("id") %>% 
  select(-c(lake, dip_other, cole_tot, hemip_tot, dip_cranefly))

env_ly = full_data %>% 
  group_by(lake,year) %>% 
  select(data_flag:DSI, -c(data_flag, block, sample_date, DOY, trap)) %>% 
  summarise(across(.fns = mean),
            n = n()) %>%  
  mutate(id = paste(lake,year, sep = "_")) %>% 
  column_to_rownames("id") %>% 
  select(-c(dailytemp, DSI, n))

#perform standardizations
sp_st = wisconsin(sqrt(sp_ly))
```

# Check univariate responses

```{r}
full_ly = bind_cols(env_ly, sp_ly)
lm1 = lm(dip_culic ~ peaktemp, data = full_data)
summary(lm1)

lm2 = lm(dip_culic ~ peaktemp, data = full_data[-c(34,123),])
summary(lm2)
#plot(lm2)

```

ephem, trich sig by fish
culic, lg chir not sig by fish

culic (+), sm dip (-) sig by peaktemp
ephem, trichop, lg chir not sig by peaktemp

lg chir sig by interaction

culic - high leverage points trap 34 and 123, lots of mosquitoes late summer in 2019 - but still sig when excluding those

# NMDS

## Choice of setup
<http://strata.uga.edu/8370/lecturenotes/multidimensionalScaling.html> Autotransform should be used for ecological data, to account for the fact that it is distributed as a count (poisson dist, use sqrt transform) and to prevent large samples from dominating the analysis (wisconsin transform).

When autotransform is not used on this dataset, dip_chir_md and dip_chir_sm dominate the axis as these are the most numerically abundant taxa.

Two and three axes both work, but with three most variation in the third axis comes from sample year so that is not informative for our questions (it compresses fish and temperature).
```{r}
#NMDSscree stuff
```

CHOICES: autotransform, bray, 2 axes, restrict set of spp




## Analysis
```{r}

mf = metaMDS(sp_ly, k=2, trace = 0)
stressplot(mf)
mf$stress

```

# Plotting

## Setup
```{r}
# setup for plotting
plot_df <- scores(mf, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("rowname") %>% 
  full_join(rownames_to_column(env_ly)) %>% 
  select(-rowname) %>% 
  mutate(fish = factor(fish))

# determine influence + significance of species
fit <- envfit(mf, sp_st, perm = 999) 

fit_pvals <- fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".")

# extract NMDS coordinates for relevant species
fit_spp <- fit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., fit_pvals, by = "species") %>% 
  filter(pvals <= 0.1) %>% 
  mutate(label = case_when(species == "dip_culic" ~ "Culicidae",
                           species == "dip_tiny" ~ "Diptera, < 1.5mm",
                           species == "ephem" ~ "Ephemeroptera",
                           species == "dip_chir_med" ~ "Chironomidae, 3-5mm",
                           species == "dip_sim_tot" ~ "Simuliidae"))

#determine influence/significance of env vars

fit2 <- envfit(mf, env_ly[,c(3,5)], perm = 999) 

fit2_pvals <- fit2$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".")

# extract NMDS coordinates for env vars
fit2_spp <- fit2 %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., fit2_pvals, by = "species") %>% 
  mutate(label = case_when(species == "fish" ~ "Fish",
                           species == "peaktemp" ~ "Peak Temperature"))

```

## Draw

```{r}
# tempcolor <- c("#FAE6CF", "#ACB945", "#008200")
custom1 <- c("#008200", "#279100", "#649F00", "#8BAD01", "#ACB945", "#C8C56C", "#DFD18F", "#F0DCB0", "#FAE6CF", "#F1F1F1")
tempcolor = c("#005200", "#008200", "#279100", "#649F00", "#8BAD01", "#ACB945", "#C8C56C", "#DFD18F", "#F0DCB0")

p = ggplot(plot_df, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = peaktemp, shape = fish), cex = 4, stroke = 2) +
  scale_shape_manual(name = "Fish", labels = c("Absent", "Present"), values = c(1,19)) +
  stat_ellipse(aes(lty= fish))+
  scale_color_gradientn(colors = rev(tempcolor),name = expression(paste("Max Temp (", degree, "C)")))+
  labs(title = "NMDS (subset) - pooled by lake/year")+
  scale_linetype_manual(values = c(3,1), guide = "none")+
  theme(plot.background = element_blank(),
        legend.key = element_rect(fill = "transparent", linetype = 0))

p + 
  geom_segment(data = fit_spp, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), lty = 5,
               col = "black") +
  geom_text_repel(data = fit_spp, aes(label = label), fontface = "italic")+
  theme(legend.position = "none")+
  geom_segment(data = fit2_spp, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "red", lwd = 1.5) +
  geom_label_repel(data = fit2_spp, aes(label = label), color = "red", segment.color = NA)+
  theme(legend.position = "right",
        plot.background = element_rect(fill = "white", color = "white"))+
  labs(title = "")

ggsave("figures/figure3.png", width = 10, height = 6, units = "in")

bind_rows(fit_spp, fit2_spp) %>% 
  select(label,pvals) %>% 
  knitr::kable()

knitr::kable(bind_rows(fit_pvals,fit2_pvals))

```

# PERMANOVA

```{r}
adonis(sp_st ~ peaktemp+fish+year, data = env_ly)

anova(betadisper(vegdist(sp_st), group = env_ly$fish))
anova(betadisper(vegdist(sp_st), group = round(env_ly$peaktemp,0)))
```

Diversity does not change with temperature or fish status; temp and fish but not their interaction are significant predictors of insect community composition.
