---
title: "slip_data_viewing"
author: "Caroline Owens"
date: "5/25/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("G:/My Drive/UCSB/Sierra/SLIP_data")
```

SLIP data is available from EDI: https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=577

```{r}
library(tidyverse)
```

```{r}
lakes = read_csv("data/Lake.csv")
lakes_sub = lakes %>% 
  select(-c(lake_drainage_name, lake_quad_name, lake_county_name, lake_juris_name, lake_wilderness_name))

frogs = read_csv("data/Amphibian.csv")

frogs_sub = frogs %>% 
  filter(amphibian_state== "Alive", amphibian_life_stage != "EggMass") %>% 
  select(-c(amphibian_location, amphibian_voucher, amphibian_state)) %>% 
  mutate(amphibian_life_stage = case_when(amphibian_life_stage == "SubAdult" ~ "Adult",
                                          T~amphibian_life_stage)) %>% 
  pivot_wider(names_from = c(amphibian_species), values_from = amphibian_number, values_fn = sum, values_fill = 0)

zCounts = read_csv("data/ZooplanktonSpeciesCounts.csv")
zNames = read_csv("data/ZooplanktonSpeciesName.csv")
zLengths = read_csv("data/ZooplanktonSpeciesLengths.csv") %>% 
  group_by(lake_id, survey_date, SpeciesID) %>% 
  summarize(length = mean(Length)) 


zoops = zCounts %>% 
  select(-Subsample) %>% 
  pivot_wider(names_from = SpeciesID, values_from = Number, values_fn = mean, values_fill = 0, names_prefix = "sp_")


```

# Figure out appropriate range of lake elevations
```{r}
lakes_sub %>% 
  filter(lake_id %in% c(50219, 50185, 50139, 50166, 51053, 51083))
```
Square - 50219

Mesa - 50185

Lower Desolation - 50139

Humphreys - 50166

No Good - 51053

Marmot - 50183


# Plot patterns of amphibian taxa over elevation
```{r}

frog_dat = frogs_sub %>% 
  left_join(lakes_sub) %>% 
  na.omit() %>% 
  filter(lake_elevation_nbr > 3000)

hyre = ggplot(frog_dat)+
  scale_x_continuous(limits = c(3000,3700))+ 
  geom_vline(xintercept = 3400)+
  labs(x = "Lake elevation (m)", y = "Hyla regilla")+
  theme(axis.title.y = element_text(face = "italic", angle = 90))+
  geom_point(aes(x=lake_elevation_nbr, y = HYRE))+
  facet_wrap(~amphibian_life_stage, scales = "free_y")#tree frog - tadpoles maybe predatory but adults disperse into terrestrial

ramu = ggplot(frog_dat)+
  scale_x_continuous(limits = c(3000,3700))+ 
  geom_vline(xintercept = 3400)+
  labs(x = "Lake elevation (m)", y = "Rana mucosa")+
  theme(axis.title.y = element_text(face = "italic", angle = 90))+
  geom_point(aes(x=lake_elevation_nbr, y = RAMU))+
  facet_wrap(~amphibian_life_stage, scales = "free_y") #MYLF

thel = ggplot(frog_dat)+
  scale_x_continuous(limits = c(3000,3700))+ 
  scale_y_continuous(limits = c(0,6))+
  geom_vline(xintercept = 3400)+
  labs(x = "Lake elevation (m)", y = "Thamnophis elegans")+
  theme(axis.title.y = element_text(face = "italic", angle = 90))+
  geom_point(aes(x=lake_elevation_nbr, y = THEL)) #garter snake

buco = ggplot(filter(frog_dat, BUCA < 10000))+
  scale_x_continuous(limits = c(3000,3700))+ 
  scale_y_log10()+
  geom_vline(xintercept = 3400)+
  labs(x = "Lake elevation (m)", y = "Bufo canorus")+
  theme(axis.title.y = element_text(face = "italic", angle = 90))+
  geom_point(aes(x=lake_elevation_nbr, y = BUCA+1))+
  facet_wrap(~amphibian_life_stage, scales = "free_y")  #yose toad - look at adults but also not as relevant as they more use meadows

```

```{r, fig.show='hold', out.width="50%"}
hyre; ramu; thel; buco
```

# Frog regressions
```{r}
hyre_A = lm(log(HYRE+1)~ lake_elevation_nbr, data = filter(frog_dat, amphibian_life_stage == "Adult" & lake_elevation_nbr >= 3400))
summary(hyre_A) #not significant

hyre_T = lm(log(HYRE+1)~ lake_elevation_nbr, data = filter(frog_dat, amphibian_life_stage == "Tadpole" & lake_elevation_nbr >= 3400))
summary(hyre_T) #significant decrease

ramu_A = lm(log(RAMU+1)~ lake_elevation_nbr, data = filter(frog_dat, amphibian_life_stage == "Adult" & lake_elevation_nbr >= 3400))
summary(ramu_A) #NS

ramu_T = lm(log(RAMU+1)~ lake_elevation_nbr, data = filter(frog_dat, amphibian_life_stage == "Tadpole" & lake_elevation_nbr >= 3400))
summary(ramu_T) #significant increase

buco_A = lm(log(BUCA+1)~ lake_elevation_nbr, data = filter(frog_dat, amphibian_life_stage == "Adult" & lake_elevation_nbr >= 3400))
summary(buco_A) #sig decrease

buco_T = lm(log(BUCA+1)~ lake_elevation_nbr, data = filter(frog_dat, amphibian_life_stage == "Tadpole" & lake_elevation_nbr >= 3400))
summary(buco_T) #sig decrease

thel_all = lm(log(THEL+1)~ lake_elevation_nbr, data = filter(frog_dat, lake_elevation_nbr >= 3400))
summary(thel_all) #NS
```

Hyla regilla tadpoles (F_(1,451) = 9.048, p = 0.00278) (a potential predator), Yosemite toad adults (F_(1,482) = 9.704, p = 0.00195) (a potential predator), and Yosemite toad tadpoles (F_(1,451) = 4.953, p = 0.0266) (a grazer) all increased at lower elevations within the range in our study lakes. Rana muscosa tadpoles decreased significantly at lower elevations within that range (F_(1,451) = 8.464, p = 0.00380) (a potential predator, particularly of smaller insects, and grazer).

# Zooplankton information

```{r}
zoop_dat = zoops %>% 
  left_join(lakes_sub) %>% 
  pivot_longer(cols = 3:40, names_prefix = "sp_", names_to = "SpeciesID") %>% 
  left_join(mutate(zNames, SpeciesID = as.character(SpeciesID)))

znl = zNames$Species_Name
names(znl) = zNames$SpeciesID

zoop_dat %>% 
  filter(lake_elevation_nbr >= 3000) %>% 
ggplot()+
  geom_point(aes(x=lake_elevation_nbr, y = value))+
  facet_wrap(~SpeciesID, scales = "free_y", labeller = as_labeller(znl))+
  scale_x_continuous(limits = c(3000,3800)) + 
  labs(x = "Lake elevation (m)", y = "Abundance")+
  geom_vline(xintercept = 3400)
```

```{r}
zoop_dat %>% 
  filter(lake_elevation_nbr >= 3400) %>% 
ggplot()+
  geom_point(aes(x=lake_elevation_nbr, y = value))+
  facet_wrap(~SpeciesID, scales = "free_y", labeller = as_labeller(znl))+
  scale_x_continuous(limits = c(3400,3800)) + 
  labs(x = "Lake elevation (m)", y = "Abundance")

zoop_dat %>% 
  filter(lake_elevation_nbr >= 3400,
         SpeciesID %in% c(1,12,16,17,18,2,3,35,36,37,40,44)) %>% #13, 7,4
  mutate(sig = case_when(SpeciesID %in% c(1,12,37,40) ~ "Yes",
                         T ~ "No")) %>% 
  ggplot(aes(x=lake_elevation_nbr, y = value))+
  geom_point()+
  facet_wrap(~SpeciesID, scales = "free_y", labeller = as_labeller(znl))+
  scale_x_continuous(limits = c(3400,3800)) + 
  labs(x = "Lake elevation (m)", y = "Abundance")+
  geom_smooth(method = "lm", formula = y~x, aes(linetype = sig), color = "skyblue")+
  scale_linetype_manual(values = c("Yes"=1, "No"=0), guide = "none")+
  #scale_y_log10()+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.line = element_line(),
        strip.text = element_text(size = 14, face = "italic"),
        axis.text = element_text(color = "black", size =12),
        axis.title = element_text(size = 18),
        plot.background = element_rect(fill = "white", colour = "white"))
ggsave("G:/My Drive/UCSB/Sierra/ch1_emergence/figures/SLIP.png", width = 13, height = 9, units = "in")

test = zoop_dat %>% 
  filter(lake_elevation_nbr >= 3400,
         SpeciesID %in% c(1,12,16,17,18,2,3,35,36,37,40,44, 13, 7, 4)) %>%#13,7,4
  dplyr::select(-c(3:5)) %>% 
  left_join(mutate(zNames, SpeciesID = as.character(SpeciesID))) %>% 
  group_by(Species_Name)

#straight LM
test1 = test %>%
  group_map(~summary(lm(value~lake_elevation_nbr, data = .)))
names(test1) = unique(test$Species_Name)
test1 = test1 %>%
  transpose()

lm_sig = tibble(unit = test1$coefficients,
               res = test1$residuals,
               Species_Name = unique(test$Species_Name))%>%
  mutate(coef = unname(map_dbl(unit, ~as_vector(.x[2,1]))),
         abs_co = abs(coef),
         pval = unname(map_dbl(unit, ~as_vector(.x[2,4]))),
         norm = unname(map_dbl(res, ~shapiro.test(.)$p.value))) %>%
  select(-unit, -res) %>%
  filter(pval <= 0.05) %>% 
  left_join(zNames)


# #log trans
# test1 = test %>% 
#   group_map(~summary(lm(log(value+1)~lake_elevation_nbr, data = .)))
# names(test1) = unique(test$Species_Name)
# test1 = test1 %>% 
#   transpose()
# 
# log_sig = tibble(unit = test1$coefficients,
#                res = test1$residuals,
#                names = unique(test$Species_Name))%>% 
#   mutate(coef = unname(map_dbl(unit, ~as_vector(.x[2,1]))),
#          abs_co = abs(coef),
#          pval = unname(map_dbl(unit, ~as_vector(.x[2,4]))),
#          norm = unname(map_dbl(res, ~shapiro.test(.)$p.value))) %>% 
#   select(-unit, -res) %>% 
#   filter(pval <= 0.05)


#square
# test1 = test %>% 
#   group_map(~summary(lm(sqrt(value)~lake_elevation_nbr, data = .)))
# names(test1) = unique(test$Species_Name)
# test1 = test1 %>% 
#   transpose()

# lsq_sig = tibble(unit = test1$coefficients,
#                res = test1$residuals,
#                names = unique(test$Species_Name))%>% 
#   mutate(coef = unname(map_dbl(unit, ~as_vector(.x[2,1]))),
#          abs_co = abs(coef),
#          pval = unname(map_dbl(unit, ~as_vector(.x[2,4]))),
#          norm = unname(map_dbl(res, ~shapiro.test(.)$p.value))) %>% 
#   select(-unit, -res) %>% 
#   filter(pval <= 0.05)
```

Daphnia melanica and dentata
Hesperodiaptomus shoshone (copepod)
Leptodiaptomus signicauda (copepod)
Keratella (?) (rotifer)

# Combined data
```{r}
dat_a = zoop_dat %>% 
  select(Species_Name, number = value, elev = lake_elevation_nbr,
         lake_id, survey_date)

dat_b = frog_dat %>% 
  select(-c(BUBO:lake_type_code, THEL)) %>% 
  pivot_longer(HYRE:BUCA, names_to = "species", values_to = "number") %>% 
  unite(Species_Name, species, amphibian_life_stage) %>%
  rename(elev = lake_elevation_nbr)

dat_all = bind_rows(dat_a, dat_b) %>% 
  filter(elev >= 3200 & elev <= 4000,
         Species_Name != "standard measurement")%>% 
  group_by(Species_Name)

#remove few observation taxa
x =dat_all %>%
  group_by(Species_Name, number != 0) %>% 
             summarise(n = n()) %>% 
  filter(`number != 0` == TRUE,
         n < 10)

dat_all = dat_all  %>%
  filter(!(Species_Name %in% x$Species_Name)) %>% 
  mutate(amphi = case_when(Species_Name %in% c("BUCA_Tadpole", "BUCA_Adult", "HYRE_Tadpole", "HYRE_Adult", "RAMU_Tadpole", "RAMU_Adult") ~ "Yes",
                           T ~ "No"))

dat = dat_all %>%
  split(.,f = .$Species_Name) %>%
  map(~summary(lm(number~elev, data = .)))%>% 
  transpose()

lm_out = tibble(unit = dat$coefficients,
                fstat = dat$fstatistic)%>%
  mutate(coef = unname(map_dbl(unit, ~as_vector(.x[2,1]))),
         abs_co = abs(coef),
         pval = unname(map_dbl(unit, ~as_vector(.x[2,4]))),
         names = names(unit),
         fst = unname(map_dbl(fstat, ~as_vector(.x[1]))),
         df = unname(map_dbl(fstat, ~as_vector(.x[3])))) %>%
  select(-unit, -fstat)# %>%
  #filter(pval <= 0.05)

holm = na.omit(lm_out) %>% 
  arrange(pval) %>% 
  mutate(j = 1:23,
         hj = 0.05/(24-j),
         condition = case_when(pval > hj ~ TRUE,
                               T ~ FALSE),
         reject_null = case_when(pval < min(pval[condition==TRUE]) ~ TRUE,
                                 T ~ FALSE)) %>% 
  filter(reject_null == TRUE) %>% 
  select(-c(j:reject_null))

lm_out = filter(lm_out,
                pval < 0.05)
```

Plot combined

```{r}
for (c in c("Yes", "No")){
p = dat_all %>%
  mutate(sig = case_when((Species_Name %in% holm$names) ~ "Yes",
                         T ~ "No")) %>% 
  filter(Species_Name %in% na.omit(lm_out)$names,
         !(Species_Name %in% c("Chaoborus trivitattus", "Daphnia pulex/pulicaria", "Simocephalus serrulatus", "nauplii")),
         amphi == c) %>% 
ggplot(aes(x=elev, y = number))+
  geom_point()+
  facet_wrap(~Species_Name, scales = "free_y")+
 # scale_x_continuous(limits = c(3200,4000)) + 
  labs(x = "Lake elevation (m)", y = "Abundance")+
  geom_smooth(method = "lm", formula = y~x, aes(linetype = sig), color = "coral2")+
  scale_linetype_manual(values = c("Yes"=1, "No"=0), guide = "none")+
  scale_y_log10()+
  theme_minimal()+
  scale_x_continuous(limits = c(3200, 3600))+
  theme(panel.grid = element_blank(),
        axis.line = element_line(),
        strip.text = element_text(size = 14, face = "italic"),
        axis.text = element_text(color = "black", size =12),
        axis.title = element_text(size = 18),
        plot.background = element_rect(fill = "white", colour = "white"),
        )
ggsave(paste0("G:/My Drive/UCSB/Sierra/ch1_emergence/figures/fig4_",c,".png"),p,
       width = 8, height = 6, units = "in")
}

```

